Clique and Constraint Models for Maximum Common (Connected) Subgraph Problems

### 2016-04-20 paper/paper.tex
Stop saying branchable

\begin{algorithm}[tb] \DontPrintSemicolon
\nl $\FuncSty{associationMCCIS}$ :: (Graph $G_1$, Graph $G_2$) $\rightarrow$ Map \;
\nl \Begin{
    \nl $\KwSty{global} ~ \mathit{incumbent} \gets \emptyset$ \label{incumbent} \;
    \nl $G \gets G_1 \modularproduct G_2$ \label{buildassoc} \;
    \nl $\FuncSty{search}(G, \emptyset, \emptyset, \operatorname{V}(G))$ \label{firstsearch} \;
    \nl $\KwSty{return} ~ \mathit{incumbent}$ \;
}
\vspace{1ex}
\nl $\FuncSty{search}$ :: (Graph $G$, Set $\mathit{solution}$, Set $\mathit{branchable}$, Set $\mathit{remaining}$) \;
\nl \Begin{
    \nl $\mathit{colourClasses} \gets \FuncSty{concatenate}($ \label{makecolours} \;
    $\hspace*{2em}\FuncSty{colour}(G, \mathit{remaining} \setminus \mathit{branchable}),$ \;
    $\hspace*{2em}\FuncSty{colour}(G, \mathit{remaining} \cap \mathit{branchable}))$ \;
    \nl \While{$\FuncSty{length}(\mathit{colourClasses}) > 0$\label{outerloop}}{
        \nl \ForEach{$v \in \FuncSty{last}(\mathit{colourClasses}) \textnormal{~in reverse order}$\label{innerloop}}{
            \nl \lIf{$|\mathit{solution}| + \FuncSty{length}(\mathit{colourClasses}) \leq
            |\mathit{incumbent}|$}{\KwSty{return}\label{bound}}
            \nl \lIf{$v \notin \mathit{branchable} ~ \KwSty{and} ~ \mathit{solution} \ne
            \emptyset$}{\KwSty{return} \label{acceptbranchable}}
            \nl $\mathit{solution}' \gets \mathit{solution} \cup \{ v \} \label{addv} \label{addvtosolution} $\;
            \nl \lIf{$|\mathit{solution}'| > |\mathit{incumbent}|$}{$\mathit{incumbent} \gets
            \mathit{solution}'$ \label{newincumbent}}
            \nl $\mathit{branchable}' \gets \mathit{branchable} \cup \{ w \in G : \FuncSty{first}(w)
            \in \operatorname{N}(G, \FuncSty{first}(v)) \}$ \label{addtobranchable} \;
            \nl $\mathit{remaining}' \gets \mathit{remaining} \cap \operatorname{N}(G, v)$ \label{filterremaining} \;
            \nl \lIf{$\mathit{remaining}' \neq \emptyset$}{$\FuncSty{search}(G, \mathit{solution}',
            \mathit{branchable}', \mathit{remaining}')$\label{recursesearch}}
         }
         \nl $\FuncSty{removeLast}(\mathit{colourClasses})$ \;
     }
}
\vspace{0.5ex}
\nl $\FuncSty{colour}$ :: (Graph $G$, Set $\mathit{uncoloured}$) $\rightarrow$ List of List of Vertex \;
\nl \Begin{
    \nl $\KwSty{return}$ a greedy sequential colouring of the vertices in $\mathit{uncoloured}$,
    using the procedure of San Segundo et al.\ \cite{DBLP:journals/cor/SegundoRJ11} with a static
    degree order from $G$ and $k_{\mathit{min}} = 0$. \;
}
\caption{An algorithm for a maximum common connected induced subgraph isomorphism via an association graph.}
\label{algorithm:mccis}
\end{algorithm}

@@

The $\mathit{branchable}$ set contains the set of association graph vertices which correspond to
vertices adjacent to an already-accepted vertex in the first input graph---in constraint programming terms, it is the
set of assignments which could be made next which maintain connectedness. (Using only one of the two
input graphs is sufficient for correctness, and has the advantage that the connectedness set
may be determined by a simple lookup into a precomputed array which maps each vertex in
the first input graph to a bitset.) At the top of search, this set is empty, and is not used (our
first vertex selection is special, and does not care about connectedness). At subsequent depths, we
may only accept vertices which are in this set, and if no such vertices remain then we return
immediately (\lineref{acceptbranchable}). When recursing, we extend $\mathit{branchable}$ with the
new vertices permitted by our acceptance of the branching $v$ (\lineref{addtobranchable}). Note that
we are assuming that inside the main loops, we encounter every vertex in $\mathit{remaining} \cap
\mathit{connected}$ before any vertex in $\mathit{remaining} \setminus \mathit{connected}$.

@@

                    \draw [color=uofgsandstone!60] (M1) -- (M2);
                    \draw [color=uofgsandstone!60] (M1) -- (M3);
                    \draw [color=uofgsandstone!60] (M2) -- (M4);
                    \draw [color=uofgsandstone!60] (M3) -- (M4);
            \end{tikzpicture}
            &
            \begin{tikzpicture}[scale=0.33]
                    \node [anchor=west] (Ns) at (0, 3.5) { $\mathit{solution}$ };
                    \node [anchor=west] (Nb) at (0, 2) { $\mathit{branchable}$ };
                    \node [anchor=west] (Nr) at (0, 0.5) { $\mathit{remaining}$ };

@@

                    \draw (Mb2) -- (Mc3);
                    \draw (Mb2) -- (Md3);
                    \draw (Mb3) -- (Mc2);
                    \draw (Mb3) -- (Md2);
            \end{tikzpicture}
            \\[0.1cm]
            \stepcounter{stepcounter2}\roman{stepcounter2}) $\mathit{remaining} \setminus \mathit{branchable}$ &
            \stepcounter{stepcounter2}\roman{stepcounter2}) $\mathit{remaining} \cap \mathit{branchable}$
            \\
        };
    \end{tikzpicture}\\[0.1cm]\begin{tikzpicture}[scale=0.33]
        \matrix (m) [matrix of nodes] {
            \begin{tikzpicture}[scale=0.33]
                \node[draw, circle, fill=chromayellow, inner sep=0.5pt, font=\normalsize] (Me4) at (0, 0) {\phantom{0}};
                \node[draw, circle, fill=chromayellow, inner sep=0.5pt, font=\normalsize] (Me5) at (3, 0) {\phantom{0}};
                \node[draw, circle, fill=chromateal, inner sep=0.5pt, font=\normalsize] (Mf4) at (6, 0) {\phantom{0}};
                \node[draw, circle, fill=chromateal, inner sep=0.5pt, font=\normalsize] (Mf5) at (9, 0) {\phantom{0}};
                \node[draw, circle, fill=chromagreen, inner sep=0.5pt, font=\normalsize] (Mb2) at (12, 0) {\phantom{0}};
                \node[draw, circle, fill=chromagreen, inner sep=0.5pt, font=\normalsize] (Mb3) at (15, 0) {\phantom{0}};
                \node[draw, circle, fill=chromanavy, inner sep=0.5pt, font=\normalsize] (Mc2) at (18, 0) {\phantom{0}};
                \node[draw, circle, fill=chromanavy, inner sep=0.5pt, font=\normalsize] (Mc3) at (21, 0) {\phantom{0}};
                \node[draw, circle, fill=chromanavy, inner sep=0.5pt, font=\normalsize] (Md2) at (24, 0) {\phantom{0}};
                \node[draw, circle, fill=chromanavy, inner sep=0.5pt, font=\normalsize] (Md3) at (27, 0) {\phantom{0}};

@@

    \caption{Solving a maximum common connected problem using an association graph. Suppose we
        have already mapped vertex $a$ to vertex $1$, giving the assignments on the right. Now we
        have two subgraphs to colour. We need two colours for $\mathit{remaining} \setminus
        \mathit{branchable}$, and we place these two colour classes first in the $\mathit{colourClasses}$
        variable. We can also colour $\mathit{remaining} \cap \mathit{branchable}$ using two
        colours, since we cannot simultaneously map $c$ to $2$ and $d$ to $3$, or vice-versa. Thus
        $\mathit{colourClasses}$ becomes a list of four colour classes. This tells us that if we hope to
        extend the current common subgraph by another four vertices, we
        must pick one assignment from each of the four colour classes (which is not actually
        possible, so the bound here gives an overestimate). The algorithm thus guesses $d \mapsto 3$ as its next
assignment, and if that fails, $d \mapsto 2$, and so on; once $b \mapsto 3$ is reached, the bound
decreases by one, and if $f \mapsto 5$ were reached we would stop due to a lack of remaining
branchable vertices. }\label{figure:assocrestricted}
\end{figure}

>>>>

\begin{algorithm}[tb] \DontPrintSemicolon
\nl $\FuncSty{associationMCCIS}$ :: (Graph $G_1$, Graph $G_2$) $\rightarrow$ Map \;
\nl \Begin{
    \nl $\KwSty{global} ~ \mathit{incumbent} \gets \emptyset$ \label{incumbent} \;
    \nl $G \gets G_1 \modularproduct G_2$ \label{buildassoc} \;
    \nl $\FuncSty{search}(G, \emptyset, \emptyset, \operatorname{V}(G))$ \label{firstsearch} \;
    \nl $\KwSty{return} ~ \mathit{incumbent}$ \;
}
\vspace{1ex}
\nl $\FuncSty{search}$ :: (Graph $G$, Set $\mathit{solution}$, Set $\mathit{connected}$, Set $\mathit{remaining}$) \;
\nl \Begin{
    \nl $\mathit{colourClasses} \gets \FuncSty{concatenate}($ \label{makecolours} \;
    $\hspace*{2em}\FuncSty{colour}(G, \mathit{remaining} \setminus \mathit{connected}),$ \;
    $\hspace*{2em}\FuncSty{colour}(G, \mathit{remaining} \cap \mathit{connected}))$ \;
    \nl \While{$\FuncSty{length}(\mathit{colourClasses}) > 0$\label{outerloop}}{
        \nl \ForEach{$v \in \FuncSty{last}(\mathit{colourClasses}) \textnormal{~in reverse order}$\label{innerloop}}{
            \nl \lIf{$|\mathit{solution}| + \FuncSty{length}(\mathit{colourClasses}) \leq
            |\mathit{incumbent}|$}{\KwSty{return}\label{bound}}
            \nl \lIf{$v \notin \mathit{connected} ~ \KwSty{and} ~ \mathit{solution} \ne
            \emptyset$}{\KwSty{return} \label{acceptbranchable}}
            \nl $\mathit{solution}' \gets \mathit{solution} \cup \{ v \} \label{addv} \label{addvtosolution} $\;
            \nl \lIf{$|\mathit{solution}'| > |\mathit{incumbent}|$}{$\mathit{incumbent} \gets
            \mathit{solution}'$ \label{newincumbent}}
            \nl $\mathit{connected}' \gets \mathit{connected} \cup \{ w \in G : \FuncSty{first}(w)
            \in \operatorname{N}(G, \FuncSty{first}(v)) \}$ \label{addtobranchable} \;
            \nl $\mathit{remaining}' \gets \mathit{remaining} \cap \operatorname{N}(G, v)$ \label{filterremaining} \;
            \nl \lIf{$\mathit{remaining}' \neq \emptyset$}{$\FuncSty{search}(G, \mathit{solution}',
            \mathit{connected}', \mathit{remaining}')$\label{recursesearch}}
         }
         \nl $\FuncSty{removeLast}(\mathit{colourClasses})$ \;
     }
}
\vspace{0.5ex}
\nl $\FuncSty{colour}$ :: (Graph $G$, Set $\mathit{uncoloured}$) $\rightarrow$ List of List of Vertex \;
\nl \Begin{
    \nl $\KwSty{return}$ a greedy sequential colouring of the vertices in $\mathit{uncoloured}$,
    using the procedure of San Segundo et al.\ \cite{DBLP:journals/cor/SegundoRJ11} with a static
    degree order from $G$ and $k_{\mathit{min}} = 0$. \;
}
\caption{An algorithm for a maximum common connected induced subgraph isomorphism via an association graph.}
\label{algorithm:mccis}
\end{algorithm}

@@

The $\mathit{connected}$ set contains the set of association graph vertices which correspond to
vertices adjacent to an already-accepted vertex in the first input graph---in constraint programming terms, it is the
set of assignments which could be made next which maintain connectedness. (Using only one of the two
input graphs is sufficient for correctness, and has the advantage that the connectedness set
may be determined by a simple lookup into a precomputed array which maps each vertex in
the first input graph to a bitset.) At the top of search, this set is empty, and is not used (our
first vertex selection is special, and does not care about connectedness). At subsequent depths, we
may only accept vertices which are in this set, and if no such vertices remain then we return
immediately (\lineref{acceptbranchable}). When recursing, we extend $\mathit{connected}$ with the
new vertices permitted by our acceptance of the branching $v$ (\lineref{addtobranchable}). Note that
we are assuming that inside the main loops, we encounter every vertex in $\mathit{remaining} \cap
\mathit{connected}$ before any vertex in $\mathit{remaining} \setminus \mathit{connected}$.

@@

                    \draw [color=uofgsandstone!60] (M1) -- (M2);
                    \draw [color=uofgsandstone!60] (M1) -- (M3);
                    \draw [color=uofgsandstone!60] (M2) -- (M4);
                    \draw [color=uofgsandstone!60] (M3) -- (M4);
            \end{tikzpicture}
            &
            \begin{tikzpicture}[scale=0.33]
                    \node [anchor=west] (Ns) at (0, 3.5) { $\mathit{solution}$ };
                    \node [anchor=west] (Nb) at (0, 2) { $\mathit{connected}$ };
                    \node [anchor=west] (Nr) at (0, 0.5) { $\mathit{remaining}$ };

@@

                    \draw (Mb2) -- (Mc3);
                    \draw (Mb2) -- (Md3);
                    \draw (Mb3) -- (Mc2);
                    \draw (Mb3) -- (Md2);
            \end{tikzpicture}
            \\[0.1cm]
            \stepcounter{stepcounter2}\roman{stepcounter2}) $\mathit{remaining} \setminus
            \mathit{connected}$ &
            \stepcounter{stepcounter2}\roman{stepcounter2}) $\mathit{remaining} \cap
            \mathit{connected}$
            \\
        };
    \end{tikzpicture}\\[0.1cm]\begin{tikzpicture}[scale=0.33]
        \matrix (m) [matrix of nodes] {
            \begin{tikzpicture}[scale=0.33]
                \node[draw, circle, fill=chromayellow, inner sep=0.5pt, font=\normalsize] (Me4) at (0, 0) {\phantom{0}};
                \node[draw, circle, fill=chromayellow, inner sep=0.5pt, font=\normalsize] (Me5) at (3, 0) {\phantom{0}};
                \node[draw, circle, fill=chromateal, inner sep=0.5pt, font=\normalsize] (Mf4) at (6, 0) {\phantom{0}};
                \node[draw, circle, fill=chromateal, inner sep=0.5pt, font=\normalsize] (Mf5) at (9, 0) {\phantom{0}};
                \node[draw, circle, fill=chromagreen, inner sep=0.5pt, font=\normalsize] (Mb2) at (12, 0) {\phantom{0}};
                \node[draw, circle, fill=chromagreen, inner sep=0.5pt, font=\normalsize] (Mb3) at (15, 0) {\phantom{0}};
                \node[draw, circle, fill=chromanavy, inner sep=0.5pt, font=\normalsize] (Mc2) at (18, 0) {\phantom{0}};
                \node[draw, circle, fill=chromanavy, inner sep=0.5pt, font=\normalsize] (Mc3) at (21, 0) {\phantom{0}};
                \node[draw, circle, fill=chromanavy, inner sep=0.5pt, font=\normalsize] (Md2) at (24, 0) {\phantom{0}};
                \node[draw, circle, fill=chromanavy, inner sep=0.5pt, font=\normalsize] (Md3) at (27, 0) {\phantom{0}};

@@

    \caption{Solving a maximum common connected problem using an association graph. Suppose we
        have already mapped vertex $a$ to vertex $1$, giving the assignments on the right. Now we
        have two subgraphs to colour. We need two colours for $\mathit{remaining} \setminus
        \mathit{connected}$, and we place these two colour classes first in the $\mathit{colourClasses}$
        variable. We can also colour $\mathit{remaining} \cap \mathit{connected}$ using two
        colours, since we cannot simultaneously map $c$ to $2$ and $d$ to $3$, or vice-versa. Thus
        $\mathit{colourClasses}$ becomes a list of four colour classes. This tells us that if we hope to
        extend the current common subgraph by another four vertices, we
        must pick one assignment from each of the four colour classes (which is not actually
        possible, so the bound here gives an overestimate). The algorithm thus guesses $d \mapsto 3$ as its next
assignment, and if that fails, $d \mapsto 2$, and so on; once $b \mapsto 3$ is reached, the bound
decreases by one, and if $f \mapsto 5$ were reached we would stop due to a lack of remaining
connected association nodes. }\label{figure:assocrestricted}
\end{figure}
